### 2025-05-01 paper/figures/style_transfer_table.tex
add paper
- Added seaborn>=0.13.2 to pyproject.toml and uv.lock for enhanced data visualization capabilities.
- Created tasks.md to outline the project checklist, covering dataset preparation, linear probing experiments, style transfer implementation, evaluation, paper writing, implementation improvements, and project management.

\begin{longtable}{cllp{5cm}p{5cm}}
\caption{Examples of Style Transfer with the Modified Phi-4 Model} \label{tab:style-transfer-examples} \\
\toprule
Example & Prompt & Formal Style Output & Informal Style Output \\
\midrule
\endfirsthead
\caption[]{Examples of Style Transfer with the Modified Phi-4 Model} \\
\toprule
Example & Prompt & Formal Style Output & Informal Style Output \\
\midrule
\endhead
\midrule
\multicolumn{4}{r}{Continued on next page} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
1 & Can you help me with this issue? &  &  \\
2 & I'd like to schedule a meeting for next week. &  Can you suggest some available time slots? &  Can you suggest a few time slots that work for you? \\
3 & The system crashed due to memory allocation failure. &  &  \\
4 & The API returns JSON data with nested objects. &  The structure is as follows:
- `data`: An array of user records.
  - Each record contains:
    - `i... &  The structure is as follows:
- `data`: An array of objects, each representing a user.
- Each user o... \\
\end{longtable}