### 2025-05-02 paper/figures/style_transfer_onehot_table.tex
paper updates

\begin{longtable}{clllcc}
\caption{Examples of Style Transfer with One-hot Style Latent Vectors} \label{tab:style-transfer-onehot} \\
\toprule
Example & Style & Prompt & Generated Text & ROUGE-1 & ROUGE-L \\
\midrule
\endfirsthead
\caption[]{Examples of Style Transfer with One-hot Style Latent Vectors} \\
\toprule
Example & Style & Prompt & Generated Text & ROUGE-1 & ROUGE-L \\
\midrule
\endhead
\midrule
\multicolumn{6}{r}{Continued on next page} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
1 & ANSWERS & The MDC has a long list of complaints it alleges are non-fulfillment of the... &  The MDC has been holding talks with the ZANU-PF led government to resolve the issues, but so far no... & 0.259000 & 0.188000 \\
1 & NEWS & The MDC has a long list of complaints it alleges are non-fulfillment of the... &  In particular, the MDC wants the government to return to the 2008 constitution which is the basis o... & 0.255000 & 0.170000 \\
2 & ANSWERS & But the political argument against focusing on the deficit is even stronger... &  They'd just see a president who had presided over a recession, and that would be that. And in the e... & 0.101000 & 0.081000 \\
2 & NEWS & But the political argument against focusing on the deficit is even stronger... &  They would be far more interested in the economy than the deficit.
That's what happened in 1993-94,... & 0.227000 & 0.124000 \\
\end{longtable}